---
title: 'Rockies Fire Analysis'
author: "Stephen Huysman"
date: "`r Sys.Date()`"
fontsize: 10pt 
output:
  html_document:
    toc: true
    toc_float: true
    self_contained: false
---
# Settings
```{r settings, include = T}
library(tidyverse)
library(multidplyr)
library(zoo)
##library(slider)
library(pROC)
library(glue)
library(parallel)
library(ggpubr)

plot_theme <- theme_bw() +
    theme(text = element_text(size=16))

#window <- 7
vars <- c("RD", "VPD", "T", "SOIL", "SWD", "RAIN", "AET", "CWD", "GDD")

### Set region to use for analysis.  Settings for regions stored in 'settings' variable
region_name <- "Middle Rockies"
region_name_clean <- str_replace(region_name, " ", "_")

settings <- tribble(
    ~region, ~params,
    "Southern Rockies", list(data_dir =  "/home/steve/OneDrive/northern-rockies-dryness/data",
                             out_dir = "/media/steve/storage/waterbalance/fire/out/ap23/",
##                                 "/home/steve/apr23_fire_points",
                             master_csv =  "SR_fire_points_extracted.csv",
                             mtbs_poly = "mtbs/Southern_rockies_fire_points.shp"
                           ),
    "Middle Rockies", list(data_dir =  "/home/steve/OneDrive/northern-rockies-dryness/data",
                           out_dir =  "/media/steve/storage/waterbalance/fire/out/",
                           master_csv =  "MR_fire_points_extracted.csv",
                           mtbs_poly = "mtbs/Middle_rockies_fire_points_v2.shp"
                           ),
)


settings <- settings %>% filter(region == region_name) 

## Directory containing metadata files
data_dir <- settings$params[[1]]$data_dir

## Directory with csv files from merge-csv.R
out_dir <- settings$params[[1]]$out_dir

master_csv <- settings$params[[1]]$master_csv
mtbs_poly <- file.path(data_dir, settings$params[[1]]$mtbs_poly)

```

```{r opts, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = paste0("/home/steve/OneDrive/northern-rockies-dryness/img/", str_replace_all(region_name, " ", "_"), "/")
)
```

# `r region_name`

# Preprocessing
1. Get MTBS database
  + Extract fire polys by location of interest (Ecoregion)
  + Create centroids for each poly (Pole of inaccessability in QGIS)
2. Get LANDFIRE EVT (Existing Vegetation Type data)
  + Use statistics tool in QGIS to get mean, median, and mode EVT by pixel count.  Only mode is used going further.
3. Create master csv file with points of interest.  Needs columns 'Event_ID', 'Incid_Type', 'Ig_Date', 'x' (=Longitude), 'y' (=Latitiude), and 'LF20_EVT_5' (Landfire 2020 EVT type).  Further analysis can look at finer resolution LANDFIRE vegetation types offered.  
4. Get Historical WB data for each centroid from Mike
5. Get Historical gridmet climate data for Each centroid
  + gridmet-historical.R - download gridmet data for each point
	+ pr, tmmx, tmmn, rmax, rmin, vpd
	+ creates one dir 'gridmet-$var' for each variable
  + merge-csv.R
    + join Mike's data and historical gridmet data into one csv for each Event_ID
	+ use these csvs for input to this script

# Master dataframe
Read in file containing meta data for fires of interest.
```{r metadata, include = T}
master <- read_csv(file.path(data_dir, master_csv)) %>%
    filter(Incid_Type == "Wildfire") %>%
##    filter(year(Ig_Date) <= 2014) %>%
    rename("maj_veg_cl" = LF20_EVT_5,
           "Acres" = BurnBndAc) %>%
    ##  maj_class is the majority
    ##    of pixels classification from Landfire 30m pixels in burned
    ##    areas, this data was extracted in QGIS.
    mutate(maj_veg_cl = case_match(maj_veg_cl,
                                   c("Herb", "Shrub", "Sparse") ~ "non_forest",
                                   "Tree" ~ "forest",
                                   c("Agriculture", "Water") ~ "other"),
           filename = paste0(Event_ID, ".csv"),
           doy = yday(Ig_Date))
```

# MTBS Stats

``` r
master %>% group_by(maj_veg_cl) %>%
  summarize(mean = mean(Acres),
            min = min(Acres),
            max = max(Acres),
            sd = sd(Acres)) %>%
  knitr::kable(format = "latex")
```

# Map

```{r map, fig.width = 7, fig.height = 5}
library(terra)
library(tidyterra)
library(maptiles)

mtbs_poly_vect <- vect(mtbs_poly) %>%
  right_join(select(master, Event_ID, maj_veg_cl),
             by = "Event_ID") %>%
  project(crs("epsg:3857"))

ecoregions <- vect("/home/steve/OneDrive/northern-rockies-dryness/data/us_eco_l3/us_eco_l3.shp") %>% project(crs("epsg:3857"))
ecoregion <- ecoregions %>% filter(US_L3NAME == region_name)

basemap <- get_tiles(buffer(ecoregion, 50000), crop = TRUE, zoom = 6, project = FALSE, provider = "CartoDB.Positron")

ggplot() +
  geom_spatraster_rgb(data = basemap, maxcell = Inf) + 
  geom_sf(data = ecoregion, inherit.aes = FALSE, alpha = 0.25, color = "midnightblue", lwd = .5) +
  geom_sf(data = mtbs_poly_vect, aes(fill = maj_veg_cl), alpha = 0.75, inherit.aes = FALSE) +
  scale_fill_manual(values = c("darkolivegreen4", "goldenrod3")) +
  ## scale_color_manual(values = c("darkolivegreen4", "khaki")) +
  labs(#x = "Longitude",
       #y = "Latitude",
       fill = "Cover") +
#       title = paste(region_name, "wildfires by cover type")) +
  theme_bw() 

summary_tbl <- master %>% group_by(maj_veg_cl) %>% summarise(n = n())
knitr::kable(summary_tbl)
```


# Determine Fire Season
```{r fire-seasons, include = T}
forest<-subset(master, maj_veg_cl=="forest")
forest_q<-quantile(forest$doy,c(0,1))#90% of data fall within this range 0.05-0.95# 0.27 to 0.75 or 0,1 to determine min max

non_forest<-subset(master, maj_veg_cl=="non_forest")
non_forest_q<-quantile(non_forest$doy,c(0,1))#90% of data fall within this range 0.05-0.95# 0.27 to 0.75 or 0,1 to determine min

#start and end dates of fire season by vegetation type derived from day of year in min_max table
doy_min_max<-as.data.frame(rbind(forest_q, non_forest_q))
forest_start<-doy_min_max[1,1]
non_forest_start<-doy_min_max[2,1]
forest_end<-doy_min_max[1,2]
non_forest_end<-doy_min_max[2,2]

knitr::kable(doy_min_max)

## Dates in data analyzed
master %>% select(Ig_Date, Acres, maj_veg_cl) %>%
    group_by(maj_veg_cl) %>%
    summarize(max_acres = max(Acres),
              min_acres = min(Acres),
              max_date = max(Ig_Date),
              min_date = min(Ig_Date)) %>%
    knitr::kable()


```

``` {r dens}
ggplot(master) +
  #geom_histogram(aes(y=..density..), colour="black", fill="white",binwidth=14) +
  geom_density(aes(x = doy, color = maj_veg_cl), alpha=.2) +
  scale_color_manual(values = c("darkolivegreen4", "goldenrod3")) +
  geom_vline(data = filter(master, maj_veg_cl == "forest"), aes(xintercept=mean(doy)), color = "darkolivegreen4", linetype="dashed", size=1) +
  geom_vline(data = filter(master, maj_veg_cl == "non_forest"), aes(xintercept=mean(doy)), color = "goldenrod3", linetype="dashed", size=1) +
  theme_bw() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  labs(title = "", x = "Day of year", y = "Fire frequency", color = NULL)+
  theme(legend.text=element_text(size=16),legend.position="right")+guides(color=guide_legend(ncol=1))

```

# Compute Percentiles
```{r percentiles, include = T}
## cores <- parallel::detectCores() - 4
## cluster <- new_cluster(cores)

my_percent_rank <- function(x) {
  ### Custom percent rank formula
  ### Implemented to fix the "pixelation" issue that was showing
  ### on maps using the NPS gridded water balance model.  Intended to
  ### reduce sensitivity to minor changes in dryness and
  ### reduce impact of large number of days during fire season with
  ### variables = 0

  ## Round to one decimal.  Fix for floating point math issues on some
  ## platforms.
  x2 <- round(x, 1)

  ## Remove zeroes.  There are many many days per fire season with
  ## variables = 0.  Causes percentiles to inflate so any day >0 is at
  ## least ~40th percentile or so, depending on the distribution.
  x2[x2 == 0] <- NA

  ## Take unique values in the time series.  Reduce impact on
  ## percentile calculation of many low dryness days.
  x2 <- unique(x2) %>% as.numeric()

  ## Ranks of unique, rounded dryness variables
  ranks <- data.frame(x = x2, pct = percent_rank(x2))

  ## Set time series of rounded rolling variables to the %ile determined
  ## on unique values above.
  inds <- match(round(x, 1), ranks$x)
  x3 <- ranks$pct[inds]

  ## zeroes were treated as NA above, replace them
  x3 <- replace_na(x3, 0)
  
  return(x3)
}

## cluster_copy(cluster, "master")
## cluster_copy(cluster, "window")
## cluster_copy(cluster, "vars")
## cluster_copy(cluster, "forest_start")
## cluster_copy(cluster, "non_forest_start")
## cluster_copy(cluster, "forest_end")
## cluster_copy(cluster, "non_forest_end")
## cluster_copy(cluster, "my_percent_rank")
## cluster_library(cluster, 'tidyverse')
## cluster_library(cluster, 'slider')
## cluster_library(cluster, 'zoo')

## files <- paste0(out_dir, master$filename)
## cluster_assign_partition(cluster, files = files)
## cluster_send(cluster, wbdata <- vroom::vroom(files, id = "PATH", altrep_opts = FALSE))
## wbdata <- party_df(cluster, "wbdata")
## master <- party_df(cluster, "master")


files <- paste0(out_dir, master$filename)
wbdata <- vroom::vroom(files, id = "PATH", altrep_opts = FALSE)

windows <- c(1, 3, 5, 7, 9, 11, 14, 17, 21, 31)

wbdata <- wbdata %>%
  mutate(Event_ID = str_extract(PATH, '(?<=/)[^/]*(?=\\.[^./]+$)')) %>% ## Extract file name without extension, use as Event_ID
  group_by(Event_ID) %>%
  mutate(RD = 100 - RH,
         WHC = max(SOIL), ## Guestimate soil WHC from max soil moisture observed
         SWD = WHC - SOIL) %>%
  left_join(select(master,
                   Event_ID, Ig_Date, maj_veg_cl, Acres),
            by = "Event_ID") %>%
  mutate(fire = if_else(Date == Ig_Date, 1, 0))  %>%
  rename(CWD = D)

percentiles <- mclapply(windows,
                        FUN = \(window) {
                          wbdata_smoothed <- select(wbdata, c(Event_ID, Ig_Date, maj_veg_cl, Acres, fire, Date, vars)) %>%
                            mutate_at(vars[1:5],
                                      ~ rollapply(.x, window, mean, by = 1, partial = FALSE, fill = NA, align = "right")) %>%
                            mutate_at(vars[6:9],
                                      ~ rollapply(.x, window, sum, by = 1, partial = FALSE, fill = NA, align = "right")) 


                          wbdata_percentiles <-
                            select(wbdata_smoothed, c(Event_ID, Ig_Date, maj_veg_cl, Acres, fire, Date, vars)) %>%
                            ## filter(Date <= Ig_Date,
                            ##        case_when(maj_veg_cl == "forest" ~
                            ##                    yday(Date) >= forest_start & yday(Date) <= forest_end,
                            ##                  maj_veg_cl == "non_forest" ~
                            ##                    yday(Date) >= non_forest_start & yday(Date) <= non_forest_end)) %>%
                            filter(Date <= Ig_Date) %>%
                            drop_na(vars) %>%
                            mutate(across(vars,
                                          my_percent_rank)) %>%
                            mutate(window = window)
                        },
                        mc.cores = 10) %>%
  bind_rows()


wbdata_percentiles_forest <- filter(percentiles, maj_veg_cl == "forest")
wbdata_percentiles_non_forest <- filter(percentiles, maj_veg_cl == "non_forest")
ignitions_forest <- wbdata_percentiles_forest %>% filter(fire == 1)
nonignitions_forest <- wbdata_percentiles_forest %>% filter(fire == 0)
ignitions_forest_long <- ignitions_forest %>% pivot_longer(cols = vars)
ignitions_non_forest <- wbdata_percentiles_non_forest %>% filter(fire == 1)
nonignitions_non_forest <- wbdata_percentiles_non_forest %>% filter(fire == 0)
ignitions_non_forest_long <- ignitions_non_forest %>% pivot_longer(cols = vars)
ignitions_long <- rbind(ignitions_forest_long, ignitions_non_forest_long)

## wbdata_smoothed_c <- wbdata_smoothed %>% collect()
## master_c <- master %>% collect()


##  percentiles %>% filter(maj_veg_cl == "forest") %>% filter(Date == "2002-06-01") %>% filter(Event_ID == "CO3701010497920020601") %>% select(D, window)
```


``` {r pct-issue, fig.width = 7, fig.height = 7}
percentiles_test <- wbdata %>%
  mutate(CWD = rollapply(CWD, 3, sum, by = 1, partial = FALSE, fill = NA, align = "right")) %>%
  mutate(CWD_orig = percent_rank(CWD),
         CWD_new = my_percent_rank(CWD)) %>%
  select(Date, Event_ID, CWD_orig, CWD_new)

ggplot(data = filter(percentiles_test, year(Date) == 2005, Event_ID == "WY4444510406019850727")) +
  geom_line(aes(x = Date, y = CWD_orig, color = "Percentile")) +
  geom_line(aes(x = Date, y = CWD_new, color = "Percentile(adj)")) +
  theme_bw() +
  labs(color = "Algorithm") +
  ylab("Percentile")

```

# Data Visualizations

```{r dates-included}
wbdata_percentiles_forest %>%
    filter(Event_ID %in% master$Event_ID[1:25]) %>%
    ggplot() +
    geom_point(aes(x = Date, y = Event_ID)) +
    geom_point(aes(x = Ig_Date, y = Event_ID), col = "red") +
    ggtitle("Dates included in analysis") +
    plot_theme

```

```{r date-distributions, eval = F, include = F}
wbdata_smoothed_c %>%
    filter(fire == 1) %>%
    pivot_longer(vars) %>%
    ggplot(mapping = aes(x = Ig_Date, y = value, color = maj_veg_cl)) +
    geom_smooth(method = "lm") +
    geom_point() +
    facet_wrap(~name, scales = "free") +
    ggtitle(paste(region_name, "water balance on ignition date"))
```

```{r eval = F}
wbdata_percentiles_forest %>%
    filter(fire == 1) %>%
    pivot_longer(vars) %>%
    ggplot(mapping = aes(x = Ig_Date, y = value, color = maj_veg_cl)) +
    geom_smooth(method = "lm") +
    geom_point() +
    facet_wrap(~name, scales = "free") +
  ggtitle(paste(region_name, "forest water balance percentile on ignition date"))

wbdata_percentiles_forest %>%
  filter(Event_ID == "UT3778610952720020601", year(Date) >= 1995, window == 7) %>%
  pivot_longer(vars) %>%
  filter(name == "CWD") %>%
  ggplot(mapping = aes(x = Date, y = value, color = maj_veg_cl)) +
  geom_line() +
  #geom_vline(xintercept = as.numeric(ymd("2002-06-01"))) +
  geom_hline(yintercept = 0.657) +
  facet_wrap(~name, scales = "free") +
  ggtitle(paste(region_name, "forest water balance percentile on ignition date"))
```

# Corrplot

```{r}
library(GGally)

ggcorr(wbdata_percentiles_forest, method = c("everything", "pearson")) 
```


<!-- ```{r} -->
<!-- wbdata_percentiles_forest %>% -->
<!--     pivot_longer(vars) %>% -->
<!--     ggplot() + -->
<!--     geom_boxplot(mapping = aes(x = name, y = value)) + -->
<!--     facet_wrap(~fire) -->


<!-- wbdata_percentiles_non_forest %>% -->
<!--     pivot_longer(vars) %>% -->
<!--     ggplot() + -->
<!--     geom_boxplot(mapping = aes(x = name, y = value)) + -->
<!--     facet_wrap(~fire) -->

<!-- wbdata_smoothed_collected <- wbdata_smoothed %>% collect() -->

<!-- wbdata_smoothed_collected %>% -->
<!--     pivot_longer(vars) %>% -->
<!--     ggplot() + -->
<!--     geom_boxplot(mapping = aes(x = name, y = value, group = fire)) -->

<!-- ``` -->

<!-- # Autocorrelation -->

<!-- ```{r} -->
<!-- wbdata %>% collect() -->

<!-- acf(wbdata %>% filter(Event_ID == "CO3942710800419870707") %>% select(P:SWD, -Event_ID)) -->
<!-- ``` -->

# ROC Analysis
We use the AUC (area under the curve) of the ROC (receiver operating characteristic) as a measure of goodness-of-fit to evaluate which predictors to use in the fire danger model.  


```{r roc_analysis, fig.width = 4, fig.height = 4, include = TRUE, eval = T, message = T, echo = T}

make_roc <- function(options) {
  roll_window = as.numeric(options[1])
  cover_class = options[2]
  
  wbdata <- percentiles %>%
    filter(window == roll_window) %>%
    filter(maj_veg_cl == cover_class)
  
  rd_roc<-roc(wbdata$fire,wbdata$RD, plot=FALSE)
  vpd_roc<-roc(wbdata$fire,wbdata$VPD, plot=FALSE)
  t_roc<-roc(wbdata$fire,wbdata$T, plot=FALSE)
  soil_roc<-roc(wbdata$fire,wbdata$SOIL, plot=FALSE)
  gdd_roc<-roc(wbdata$fire,wbdata$GDD, plot=FALSE)
  swd_roc<-roc(wbdata$fire,wbdata$SWD, plot=FALSE)
  rain_roc<-roc(wbdata$fire,wbdata$RAIN, plot=FALSE)
  aet_roc<-roc(wbdata$fire,wbdata$AET, plot=FALSE)
  cwd_roc<-roc(wbdata$fire,wbdata$CWD, plot=FALSE)

  png(filename = paste0("/home/steve/OneDrive/northern-rockies-dryness/img/", region_name_clean, "/", region_name, "-", cover_class, "-", roll_window, "-days.png"), width = 600, height = 600)
  par(pty = "s") ## pty sets the aspect ratio of the plot region. Two options:
  ##                "s" - creates a square plotting region
  ##                "m" - (the default) creates a maximal plotting region
  par(bg="white")
  par(cex.axis = 1.75)
  par(cex.lab = 2)
  print({
    plot.roc(col="pink",rd_roc, lwd=4, print.auc=FALSE, legacy.axes = TRUE, xlab="False Positive Percentage", ylab="True Postive Percentage")#, main = paste(region_name, cover_class, roll_window, "days"))
    plot.roc(add=TRUE,col="dark blue",vpd_roc, main="vpd", lwd=4, print.auc=FALSE)
    plot.roc(add=TRUE,col="green",t_roc, main="temp", lwd=4, print.auc=FALSE)
    plot.roc(add=TRUE,col="red",soil_roc, main="soil", lwd=4, print.auc=FALSE)
    plot.roc(add=TRUE,col="brown",gdd_roc, main="gdd", lwd=4, print.auc=FALSE)
    plot.roc(add=TRUE,col="magenta",swd_roc, main="swd", lwd=4, print.auc=FALSE)
    plot.roc(add=TRUE,col="purple",rain_roc, main="rain", lwd=4, print.auc=FALSE)
    plot.roc(add=TRUE,col="dark green",aet_roc, main="aet", lwd=4, print.auc=FALSE)
    plot.roc(add=TRUE,col="orange",cwd_roc, main="cwd", lwd=4, print.auc=FALSE)
    # Add a legend
    legend(.35, .75, legend=c(paste("RD ",round(auc(rd_roc),2)),
                              #paste("SVP ",round(auc(svp_roc),2)),
                              paste("VPD ",round(auc(vpd_roc),2)),
                              paste("T ",round(auc(t_roc),2)),
                              paste("SOIL ",round(auc(soil_roc),2)),
                              paste("GDD ",round(auc(gdd_roc),2)),
                              paste("SWD",round(auc(swd_roc),2)),
                              paste("RAIN ",round(auc(rain_roc),2)),
                              paste("AET ",round(auc(aet_roc),2)),
                              paste("CWD ",round(auc(cwd_roc),2))),
           bty="n", col=c("pink", "dark blue","green","red","brown","magenta","purple","dark green","orange"), #col=c("black","blue", "dark blue","green","red","brown","magenta","purple","dark green","orange")
           lty=1, lwd=4, text.font=2, cex=1.5)
  })
  dev.off()

  auc0 <- function(var) {
    var <- tolower(var)
    roc <- paste0(var, "_roc")
    return(as.numeric(auc(get(roc))))
  }

  auc0lb <- function(var) {
    ### Lower-bound estimate for full AUC
    var <- tolower(var)
    roc <- paste0(var, "_roc")
    return(as.numeric(ci(get(roc)))[1])
  }

  auc0ub <- function(var) {
    ### Upper-bound estimate for full AUC
    var <- tolower(var)
    roc <- paste0(var, "_roc")
    return(as.numeric(ci(get(roc)))[3])
  }
  
  auc10 <- function(var) {
    var <- tolower(var)
    roc <- paste0(var, "_roc")
    return(as.numeric(auc(get(roc),  partial.auc=c(1, .9))))
  }

  auc20 <- function(var) {
    var <- tolower(var)
    roc <- paste0(var, "_roc")
    return(as.numeric(auc(get(roc),  partial.auc=c(1, .8))))
  }

  roc_table <- tribble(~window, ~maj_veg_cl, ~var, ~auc, ~aucub, ~auclb, ~auc20, ~auc10)
  for (var in vars) {
    roc_table <- roc_table %>% add_row(window = roll_window, maj_veg_cl = cover_class, var = var, auc = auc0(var), auclb = auc0lb(var), aucub = auc0ub(var), auc20 = auc20(var), auc10 = auc10(var))
  }

  return(roc_table)
}

roc_options <- expand.grid(
  window = windows,
  maj_veg_cl = c("forest", "non_forest")
) %>%
  t() %>%
  data.frame()

roc_table <- mclapply(roc_options, make_roc, mc.cores = 10) %>%
  bind_rows()


### Top predictors Full AUC
### Forest
roc_table %>%
  filter(maj_veg_cl == "forest") %>%
  arrange(desc(auc)) %>%
  head(n = 10) %>%
  knitr::kable()

### Non_forest
roc_table %>%
  filter(maj_veg_cl == "non_forest") %>%
  arrange(desc(auc)) %>%
  head(n = 10) %>%
  knitr::kable()

### Top predictors AUC20
### Forest
roc_table %>%
  filter(maj_veg_cl == "forest") %>%
  arrange(desc(auc20))%>%
  head(n = 10) %>%
  knitr::kable()

### Non_forest
roc_table %>%
  filter(maj_veg_cl == "non_forest") %>%
  arrange(desc(auc20))%>%
  head(n = 10) %>%
  knitr::kable()

### Top predictors AUC10
### Forest
roc_table %>%
  filter(maj_veg_cl == "forest") %>%
  arrange(desc(auc10))%>%
  head(n = 10) %>%
  knitr::kable()

### Non_forest
roc_table %>%
  filter(maj_veg_cl == "non_forest") %>%
  arrange(desc(auc10))%>%
  head(n = 10) %>%
  knitr::kable()
```

### Window vs AUC
``` {r auc-window, fig.height = 8, fig.width = 6}
forest_auc_window <- roc_table %>%
  filter(maj_veg_cl == "forest") %>%
  ggplot() +
  geom_line(aes(x = window, y = auc, color = var)) +
  geom_ribbon(aes(x = window, ymin = auclb, ymax = aucub, color = var), linetype = 2, alpha = 0.05) +
  labs(title = "Forest AUC") +
  theme_bw()

non_forest_auc_window <- roc_table %>%
  filter(maj_veg_cl == "non_forest") %>%
  ggplot() +
  geom_line(aes(x = window, y = auc, color = var)) +
  geom_ribbon(aes(x = window, ymin = auclb, ymax = aucub, color = var), linetype = 2, alpha = 0.05) +
  labs(title = "Non_forest AUC") +
  theme_bw()

forest_pauc20_window <- roc_table %>%
  filter(maj_veg_cl == "forest") %>%
  ggplot() +
  geom_line(aes(x = window, y = auc20, color = var)) +
  labs(title = "Forest pAUC (.2)") +
  theme_bw()

non_forest_pauc20_window <- roc_table %>%
  filter(maj_veg_cl == "non_forest") %>%
  ggplot() +
  geom_line(aes(x = window, y = auc20, color = var)) +
  labs(title = "Non_forest pAUC (.2)") +
  theme_bw()

forest_pauc10_window <- roc_table %>%
  filter(maj_veg_cl == "forest") %>%
  ggplot() +
  geom_line(aes(x = window, y = auc10, color = var)) +
  labs(title = "Forest pAUC (.1)") +
  theme_bw()

non_forest_pauc10_window <- roc_table %>%
  filter(maj_veg_cl == "non_forest") %>%
  ggplot() +
  geom_line(aes(x = window, y = auc10, color = var)) +
  labs(title = "Non_forest pAUC (.1)") +
  theme_bw()

window_plot <- ggarrange(forest_auc_window, non_forest_auc_window,
                         forest_pauc20_window, non_forest_pauc20_window,
                         forest_pauc10_window, non_forest_pauc10_window,
                         ncol = 2,
                         nrow = 3,
                         legend = "right",
                         common.legend = TRUE
                         )

annotate_figure(window_plot)

```


demonstrate histogram of conditions associated with fire or no fire to see how much overlap there is.  This helps interpret the ROC
```{r density-hist, echo = F, message = F}
plot(density(ignitions_forest$CWD), col="red")
lines(density(nonignitions_forest$CWD), col="blue")

plot(density(ignitions_forest$VPD), col="red")
lines(density(nonignitions_forest$VPD), col="blue")
```

<!-- Determine contribution of various parameters to burned area -->
<!-- ```{r burned-area-contrib, echo = F, message = F, eval = F} -->
<!-- library(hier.part) -->

<!-- y<-as.vector(ignitions_non_forest[,"Acres"]) -->
<!-- xcann<-subset(ignitions_non_forest,select=c("T","SOIL","GDD","SWD","AET","CWD","RD","VPD")) -->
<!-- hier.part(y,xcann,barplot=TRUE) -->

<!-- y<-as.vector(ignitions_forest[,"Acres"]) -->
<!-- xcann<-subset(ignitions_forest,select=c("T","SOIL","GDD","SWD","AET","CWD","RD","VPD")) -->
<!-- hier.part(y,xcann,barplot=TRUE) -->
<!-- ``` -->

Plots of wb percentile vs. area burned show weak relationships
```{r wb-vs-area, fig.width = 5, fig.height = 5, eval = F, include = F}
ggplot(ignitions_long, aes(x=value, y=Acres, colour=factor(maj_veg_cl))) +
    geom_point(size=3) +
    geom_smooth(method="lm", se=FALSE) +
    facet_wrap(~name, scales="fixed") +
    labs(title = "", x = "Percentile at day of ignition", y = "Area burned (Acres)", color = NULL)+
    scale_color_manual(values=c("red","black")) +
    plot_theme
```

```{r wb-vs-logarea, fig.width = 8, fig.height = 8, echo = F, message = F}
ggplot(ignitions_long, aes(x=value, y=log(Acres), colour=factor(maj_veg_cl))) +
    geom_point(size=3, alpha = 0.10) +
    geom_smooth(method="lm", se=FALSE, formula = y ~ poly(x, 2)) +
    facet_wrap(~name, scales="fixed") +
    labs(title = "", x = "Percentile at day of ignition", y = "Area burned (log Acres)", color = NULL)+
    scale_color_manual(values=c("red","black")) +
    plot_theme
```

# ECDF

```{r ecdf-all-vars, fig.width = 8, fig.height = 4, echo = F}
ggplot(ignitions_long, aes(value, color = name)) +
    stat_ecdf(pad = FALSE) +
    xlab("Percentile at ignition") +
    facet_grid(~ maj_veg_cl) +
    plot_theme
```

```{r ecdf-cwd-vpd, fig.width = 8, fig.height = 4, echo = F} 
ggplot(filter(ignitions_long,
              name %in% c("VPD", "CWD")),
       aes(value, color = name)) +
    stat_ecdf(pad = FALSE) +
    geom_hline(yintercept = .1) +
    geom_hline(yintercept = .4) +
    xlab("Percentile at ignition") +
    ylab("Proportion of Fires") +
    facet_grid(~ maj_veg_cl) +
    plot_theme
```

# Fire Danger Model

The fire danger model uses ECDF curves of water balance percentiles on day of ignition.  Water balance variables are selected based on AUC values in the ROC analysis above. We can use the logistic regression models evaluated in the ROC analysis to estimate probability of ignition for given water balance percentiles for a given day, but due to the very large number of non-ignition days in the dataset, the probability of ignition for any given day will be extremely low.  Instead, we will look only at days that had an ignition, and compare the water balance percentiles for those days.  This allows us to, for example, take a percentile of water deficit on a given day and compare that to the proportion of historic fires that burned at or below this value.  

Low fire danger is taken as the water balance percentile at or below which 10% of historical fires ignited.

High fire danger is taken as the water balance percentile at or below which 40% of historical fires 

A regression line is fit to the ECDF curves to develop an equation that can be used to estimate proportion of historical fires for a given cover type burning at or below an input water balance variable percentile. 

```{r ecdf-regression-setup}
selected_vars <- c("RD", "VPD", "CWD")

regression_formulas <- tribble(
    ~cover, ~var, ~order, ~formula, ~r2
)

precision <- 7

regression1_formula <- function(a, b) {
    ## Regression formula in string format
    a <- round(a, precision)
    b <- round(b, precision)
    return (paste0("y = ", a, "e^(", b, "x)"))
}

regression2_formula <- function(b0, b1, b2) {
    ## Regression formula in string format
    b0 <- round(b0, precision)
    b1 <- round(b1, precision)
    b2 <- round(b2, precision)
    return (paste0("y = e^(", b0, " + ", b1, "x + ", b2, "x^2)"))
}

regression3_formula <- function(b0, b1, b2, b3) {
    ## Regression formula in string format
    b0 <- round(b0, precision)
    b1 <- round(b1, precision)
    b2 <- round(b2, precision)
    b3 <- round(b3, precision)
    return (paste0("y = e^(", b0, " + ", b1, "x + ", b2, "x^2 + ", b3, "x^3)"))
}

### Selected window of time to view eCDF curves
ecdf_window <- 7

forest_data <- ignitions_forest %>% filter(window == ecdf_window)
non_forest_data <- ignitions_non_forest %>% filter(window == ecdf_window)
```


## Third Order

### CWD Forest

```{r ecdf-regression3-d-forest}
ecdf_fn <- ecdf(forest_data$CWD)
summary(ecdf_fn)

estimated <- data.frame(
    dec = seq(from = 0.01, to = 0.99, by = 0.01)
)

estimated$quants <- quantile(ecdf_fn, estimated$dec)
estimated$proportions <- ecdf_fn(estimated$quants)

lm <- lm(log(proportions) ~ poly(quants, 3, raw = TRUE), data = estimated)
r2 <- summary(lm)$adj.r.squared
estimated$predicted <- exp(predict(lm, data = estimated$quants))

##plot(lm)

summary(lm)

b0 <- lm$coefficients[[1]]
b1 <- lm$coefficients[[2]]
b2 <- lm$coefficients[[3]]
b3 <- lm$coefficients[[4]]

threshold <- 0.35 ### Threshold deficit %ile for low fire danger

plot(ecdf_fn,
     xlim = c(0, 1),
     ylim = c(0, 1),
     main = "",
     xlab = "Percentile deficit",
     ylab = "Historical proportion of fires",
     cex.lab = 1.25)
#points(y = estimated$proportions, x = estimated$quants, col = "blue")
points(y = estimated$predicted, x = estimated$quants, col = "red")
curve(exp(b0 + b1*x + b2*(x^2) + b3*(x^3)), add = TRUE, col = "red")
### Add example of low fire danger rating
y1 <- exp(predict(lm, list(quants = c(0.35))))
segments(-.1, y1, threshold, y1)
segments(threshold, y1, threshold, -.1)
mtext(round(y1, 2), 2, at=y1, col=2, cex = 1.25)
mtext(threshold, 1, at=threshold, col=2, cex = 1.25)

##curve(exp(predict(lm, newdata = data.frame(quants = x))), add = TRUE, col = "red")



regression_formulas <- regression_formulas %>%
    add_row(var = "CWD", order = "3", cover = "forest", formula = regression3_formula(b0, b1, b2, b3), r2 = r2)
```

Estimated function predicting proportion of historic fires burning in forests at or below given percentile of CWD on day of ignition:

$\begin{align*}f(x) = e^{`r b0` + `r b1`x + `r b2`x^2 + `r b3`x^3}\end{align*}$

### VPD Forest

```{r ecdf-regression3-vpd-forest}
ecdf_fn <- ecdf(forest_data$VPD)
summary(ecdf_fn)

estimated <- data.frame(
    dec = seq(from = 0.01, to = 0.99, by = 0.01)
)

estimated$quants <- quantile(ecdf_fn, estimated$dec)
estimated$proportions <- ecdf_fn(estimated$quants)

lm <- lm(log(proportions) ~ poly(quants, 3, raw = TRUE), data = estimated)
r2 <- summary(lm)$adj.r.squared
estimated$predicted <- exp(predict(lm, data = estimated$quants))

##plot(lm)

summary(lm)

b0 <- lm$coefficients[[1]]
b1 <- lm$coefficients[[2]]
b2 <- lm$coefficients[[3]]
b3 <- lm$coefficients[[4]]

plot(ecdf_fn)
points(y = estimated$proportions, x = estimated$quants, col = "blue")
points(y = estimated$predicted, x = estimated$quants, col = "red")
curve(exp(b0 + b1*x + b2*(x^2) + b3*(x^3)), add = TRUE, col = "red")
##curve(exp(predict(lm, newdata = data.frame(quants = x))), add = TRUE, col = "red")

regression_formulas <- regression_formulas %>%
    add_row(var = "VPD", order = "3", cover = "forest", formula = regression3_formula(b0, b1, b2, b3), r2 = r2)
```

Estimated function predicting proportion of historic fires burning in forests at or below given percentile of VPD on day of ignition:

$\begin{align*}f(x) = e^{`r b0` + `r b1`x + `r b2`x^2 + `r b3`x^3}\end{align*}$

### CWD Non-forest

```{r ecdf-regression3-d-non_forest}
ecdf_fn <- ecdf(non_forest_data$CWD)
summary(ecdf_fn)

estimated <- data.frame(
    dec = seq(from = 0.01, to = 0.99, by = 0.01)
)

estimated$quants <- quantile(ecdf_fn, estimated$dec)
estimated$proportions <- ecdf_fn(estimated$quants)

lm <- lm(log(proportions) ~ poly(quants, 3, raw = TRUE), data = estimated)
r2 <- summary(lm)$adj.r.squared
estimated$predicted <- exp(predict(lm, data = estimated$quants))

##plot(lm)

summary(lm)

b0 <- lm$coefficients[[1]]
b1 <- lm$coefficients[[2]]
b2 <- lm$coefficients[[3]]
b3 <- lm$coefficients[[4]]

plot(ecdf_fn, xlab = "Percentile of Deficit", ylab = "Proportion of historical wildfires", main = "eCDF Non_Forest Deficit, with ignition danger ratings", xlim = c(0,1))
points(y = estimated$proportions, x = estimated$quants, col = "blue")
points(y = estimated$predicted, x = estimated$quants, col = "red")
curve(exp(b0 + b1*x + b2*(x^2) + b3*(x^3)), add = TRUE, col = "red")
##curve(exp(predict(lm, newdata = data.frame(quants = x))), add = TRUE, col = "red")
## low_thresh <- .1
## med_thresh <- .4
## high_thresh <- .7
## x_low <- .46
## segments(-5, low_thresh, x1 = x_low, y1 = low_thresh) ## Low Fire Danger horiz
## text(x = .15, y = low_thresh + 0.02, label = "Low Fire Danger")
## segments(x_low, -5, x1 = x_low, y1 = low_thresh) ## Low Fire Danger vert
## text(x = x_low + .025, y = -.02, label = x_low)
## x_med <- .87
## segments(-5, med_thresh, x1 = x_med, y1 = med_thresh) ## High Fire Danger horiz
## text(x = .15, y = med_thresh + 0.02, label = "Medium Fire Danger")
## segments(x_med, -5, x1 = x_med, y1 = med_thresh) ## High Fire Danger vert
## text(x = x_med + .025, y = -.02, label = x_med)
## x_high <- .96
## segments(-5, high_thresh, x1 = x_high, y1 = high_thresh) ## High Fire Danger horiz
## text(x = .15, y = high_thresh + 0.02, label = "High Fire Danger")
## segments(x_med, -5, x1 = x_high, y1 = high_thresh) ## High Fire Danger vert
## text(x = x_high + .025, y = -.02, label = x_high)



regression_formulas <- regression_formulas %>%
    add_row(var = "CWD", order = "3", cover = "non_forest", formula = regression3_formula(b0, b1, b2, b3), r2 = r2)
```

Estimated function predicting proportion of historic fires burning in non_forests at or below given percentile of CWD on day of ignition:

$\begin{align*}f(x) = e^{`r b0` + `r b1`x + `r b2`x^2 + `r b3`x^3}\end{align*}$

### VPD Non-forest

```{r ecdf-regression3-vpd-non_forest}
ecdf_fn <- ecdf(non_forest_data$VPD)
summary(ecdf_fn)

estimated <- data.frame(
    dec = seq(from = 0.01, to = 0.99, by = 0.01)
)

estimated$quants <- quantile(ecdf_fn, estimated$dec)
estimated$proportions <- ecdf_fn(estimated$quants)

lm <- lm(log(proportions) ~ poly(quants, 3, raw = TRUE), data = estimated)
r2 <- summary(lm)$adj.r.squared
estimated$predicted <- exp(predict(lm, data = estimated$quants))

##plot(lm)

summary(lm)

b0 <- lm$coefficients[[1]]
b1 <- lm$coefficients[[2]]
b2 <- lm$coefficients[[3]]
b3 <- lm$coefficients[[4]]

solve_ecdf <- function(x) {
    exp(b0 + b1*x + b2*(x^2) + b3*(x^3))
}

plot(ecdf_fn)
points(y = estimated$proportions, x = estimated$quants, col = "blue")
points(y = estimated$predicted, x = estimated$quants, col = "red")
curve(exp(b0 + b1*x + b2*(x^2) + b3*(x^3)), add = TRUE, col = "red")
## ##curve(exp(predict(lm, newdata = data.frame(quants = x))), add = TRUE, col = "red")
## x_low <- round(log(low_thresh/a)/b,2)
## segments(0, low_thresh, x1 = x_low, y1 = low_thresh) ## Low Fire Danger horiz
## text(x = .3, y = low_thresh + 0.02, label = "Low Fire Danger")
## segments(x_low, 0, x1 = x_low, y1 = low_thresh) ## Low Fire Danger vert
## text(x = x_low, y = -.02, label = x_low)
## x_high <- round(log(high_thresh/a)/b,2)
## segments(0, high_thresh, x1 = x_high, y1 = high_thresh) ## High Fire Danger horiz
## text(x = .3, y = high_thresh + 0.02, label = "High Fire Danger")
## segments(x_high, 0, x1 = x_high, y1 = high_thresh) ## High Fire Danger vert
## text(x = x_high, y = -.02, label = x_high)

regression_formulas <- regression_formulas %>%
    add_row(var = "VPD", order = "3", cover = "non_forest", formula = regression3_formula(b0, b1, b2, b3), r2 = r2)
```

Estimated function predicting proportion of historic fires burning in non_forests at or below given percentile of VPD on day of ignition:

$\begin{align*}f(x) = e^{`r b0` + `r b1`x + `r b2`x^2 + `r b3`x^3}\end{align*}$

### RD Non-forest

```{r ecdf-regression3-rd-non_forest}
ecdf_fn <- ecdf(non_forest_data$RD)
summary(ecdf_fn)

estimated <- data.frame(
    dec = seq(from = 0.01, to = 0.99, by = 0.01)
)

estimated$quants <- quantile(ecdf_fn, estimated$dec)
estimated$proportions <- ecdf_fn(estimated$quants)

lm <- lm(log(proportions) ~ poly(quants, 3, raw = TRUE), data = estimated)
r2 <- summary(lm)$adj.r.squared
estimated$predicted <- exp(predict(lm, data = estimated$quants))

##plot(lm)

summary(lm)

b0 <- lm$coefficients[[1]]
b1 <- lm$coefficients[[2]]
b2 <- lm$coefficients[[3]]
b3 <- lm$coefficients[[4]]

plot(ecdf_fn)
points(y = estimated$proportions, x = estimated$quants, col = "blue")
points(y = estimated$predicted, x = estimated$quants, col = "red")
curve(exp(b0 + b1*x + b2*(x^2) + b3*(x^3)), add = TRUE, col = "red")
##curve(exp(predict(lm, newdata = data.frame(quants = x))), add = TRUE, col = "red")

regression_formulas <- regression_formulas %>%
    add_row(var = "RD", order = "3", cover = "non_forest", formula = regression3_formula(b0, b1, b2, b3), r2 = r2)
```

Estimated function predicting proportion of historic fires burning in non_forests at or below given percentile of  `window` day rolling mean of RD on day of ignition:

$\begin{align*}f(x) = e^{`r b0` + `r b1`x + `r b2`x^2 + `r b3`x^3}\end{align*}$

## Estimated Regression Formulas
`r knitr::kable(regression_formulas)`


# Dryness vs wildfire

``` {r forest-dens, fig.height = 6, fig.width = 6}
wbdata_percentiles_forest %>%
  filter(window == 3) %>%
  mutate(fire = factor(fire)) %>%
  pivot_longer(cols = c(RD, VPD, T, SOIL, SWD, RAIN, AET, CWD, GDD)) %>%
  ggplot() +
  geom_density(aes(value, y = ..scaled.., fill = fire), alpha = 0.5) +
  scale_fill_manual(values = c("blue","red")) +
  labs(x = "Percentile", y = "Density") +
  facet_wrap(~name, scales = "free") 
```

``` {r non_forest-dens, fig.height = 6, fig.width = 6}
wbdata_percentiles_non_forest %>%
  filter(window == 3) %>%
  mutate(fire = factor(fire)) %>%
  pivot_longer(cols = c(RD, VPD, T, SOIL, SWD, RAIN, AET, CWD, GDD)) %>%
  ggplot() +
  geom_density(aes(value, y = ..scaled.., fill = fire), alpha = 0.5) +
  scale_fill_manual(values = c("blue","red")) +
  labs(x = "Percentile", y = "Density") +
  facet_wrap(~name, scales = "free") 
```


# R Version

```{r session-info}
sessionInfo()
```
